---
title: "Exploring climate data"
author: "Nick Lauerman"
date: "February 11, 2019"
output: 
  pdf_document: 
    fig_caption: yes
    fig_height: 6.5
    fig_width: 6.5
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# load data
The Waukegan Regional Airport data will be used as the test data set for
developing the load, or import, process. Remove data that is standard in the
download.
```{r}
KUGN <- read.csv(file = "./data/Waukegan/obs.csv",
                 stringsAsFactors = FALSE)
KUGN$STATION <- "Waukagen Airport"
KUGN$NAME <- NULL
```

# Format data
Format the date froma text field to a date data type and compute the Average Temperature (TAVG). This is not always computed for a station.
```{r}
KUGN$DATE <- as.Date(KUGN$DATE,
                     format = "%Y-%m-%d")
KUGN$TAVG <- (KUGN$TMAX + KUGN$TMIN)/2 
```

# Set variable names
Change variable names to names that comply with the coding standard
```{r}
Cnames <- c("Station",
            "Date",
           "Precip",
           "AverageTemp",
           "MaxTemp",
           "MinTemp")
names(KUGN) <- Cnames
rm(Cnames)
```

#norms
```{r}
kugnNorms <- read.csv("./data/Waukegan/norms.csv",
                      stringsAsFactors = FALSE)
kugnNorms$STATION <- "Waukagen Airport"
kugnNorms$STATION_NAME <- NULL
kugnNorms$DATE <- as.Date(as.character( kugnNorms$DATE),
                          format = "%Y%m%d")
Cnames <- c("Station",
            "Date",
            "averageTemp",
            "averageDutrTemp",
            "normalMaxTemp",
            "normalMinTemp",
            "averageTempSD",
            "averageDutrTempSD",
            "maxTempSD",
            "minTempSD")
names(kugnNorms) <- Cnames
rm(Cnames)
```




#graphical exploration
Couple of plots to check data
```{r}
plot(x = KUGN$Date, y = KUGN$MaxTemp,
     type = "l",
     col = "red",
     main = "Daily Max Tempature",
     xlab = "Day",
     ylab = "tempature (F)")
plot(x = KUGN$Date, y=KUGN$MinTemp,
     type = "l",
     col = "blue",
     main = "Daily Min Tempature",
     xlab = "Day",
     ylab = "tempature (F)")
plot(x = KUGN$Date, y = KUGN$Precip,
     type = "l",
     col = "red",
     main = "Daily precip",
     xlab = "Day",
     ylab = "inches")
plot(x= kugnNorms$Date, y=kugnNorms$normalMaxTemp,
     type = "l",
     col = "red",
     main = "Normal Daily Max Tempature",
     xlab = "Day",
     ylab = "tempature (F)")
plot(x= kugnNorms$Date, y=kugnNorms$normalMinTemp,
     type = "l",
     col = "blue",
     main = "Normal Daily Min Tempature",
     xlab = "Day",
     ylab = "tempature (F)")


```


